Wireless Forensics Principles and Investigations

### 2021-11-24 paper.tex
Added evans section, fix pics

%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[acmlarge]{acmart}

\usepackage{lipsum}

@@

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
 \setcopyright{acmcopyright}
 \copyrightyear{2021}
 \acmYear{2021}
 \acmDOI{}

@@

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

@@

%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Lomax, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
We will discuss the current and emerging principles and investigations as it pertains to wireless digital forensics. We will look at the past, the present and future of wireless forensics. We will examine the major sub-branches, to include: 802.11 network forensics, IoT vulnerabilities, mobile forensics, available wireless security and forensics software tools. We will compare and contrast some of the major forensic tool suites as well as some emerging technologies. 
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10003033.10003083.10003014</concept_id>
       <concept_desc>Networks~Network security</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10002978.10003014.10003017</concept_id>
       <concept_desc>Security and privacy~Mobile and wireless security</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10002978.10002979.10002983</concept_id>
       <concept_desc>Security and privacy~Cryptanalysis and other attacks</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

@@

\section{Introduction}
%% TODO add introduction

@@

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

>>>>

\documentclass[acmlarge]{style/acmart}

\usepackage{lipsum}
\usepackage{float}

@@

 \setcopyright{acmcopyright}
 \copyrightyear{2021}
 \acmYear{2021}
 \acmDOI{}

@@

@@

\renewcommand{\shortauthors}{Lomax, et al.}

\begin{abstract}
We will discuss the current and emerging principles and investigations as it pertains to wireless digital forensics. We will look at the past, the present and future of wireless forensics. We will examine the major sub-branches, to include: 802.11 network forensics, IoT vulnerabilities, mobile forensics, available wireless security and forensics software tools. We will compare and contrast some of the major forensic tool suites as well as some emerging technologies.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10003033.10003083.10003014</concept_id>
       <concept_desc>Networks~Network security</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10002978.10003014.10003017</concept_id>
       <concept_desc>Security and privacy~Mobile and wireless security</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10002978.10002979.10002983</concept_id>
       <concept_desc>Security and privacy~Cryptanalysis and other attacks</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

@@

\section{Introduction}
%% TODO add more introduction

@@

\subsection{Best Practices}

As demonstrated from above

\section{802.11 and Mobile Forensic Tools}

802.11 or “WiFi” has become normal not only in the workplace but also at home. Because of the common use of this technology, it has become necessary to develop tools and techniques in order to conduct forensic investigations on WiFi networks. These tools can simply be used to monitor traffic to find out why a network may be running slower than normal or they can be utilized for in-depth investigations because of an attack. There are also several tools and techniques for attackers or hackers to utilize in order to inject malicious packets, intercept packets, and crack network passwords in order to gain unauthorized access to a wireless network. 
There are several avenues for an examiner to conduct a forensic investigation into a wireless network. The first is examining frames on the network and capturing network traffic or packets utilizing tools and hardware. This will allow the examiner to inspect communications between stations. Packets can also contain clear-text data if unencrypted or if the encryption key is known. Most WiFi networks will typically operate in either 2.4GHz or 5 GHz radio frequency ranges. WiFi ranges are broken down into smaller bands called channels. 2.4GHz band will have 11 channels and 5GHz band will have 45 different channels. 
Wireless Access Points (WAP) can also contain fruitful information for a forensic examiner. WAPs can contain stored logs which will illuminate connections whether they were successful or not. The logs could also contain data about the attempted or successful connection such as IP and MAC addresses as well as timestamps. 
%Below is a screenshot example of a wireless log from a WAP:

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\linewidth]{imgs/waplog.jpg}
  \caption{Screenshot of a wireless log from a WAP}
  \label{fig:waplog}
\end{figure}

This particular log shows real-time connections which can be volatile. Forensic examiners can use real-time logs to determine if an unauthorized client is connected but historic logs can obviously be extremely useful as well.

\subsection{Wireshark}

In order to an examiner to capture network traffic, an 802.11 wireless card with “monitor mode” capability is required. Wireshark is a useful tool and probably the most used by forensic examiners for monitoring network traffic live and examining historic pcap files. Wireshark allows the examiner to analyze traffic and filter exactly what they are looking for. 
%Below is a screen shot of a live capture using Wireshark:

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\linewidth]{imgs/wireshark.jpg}
  \caption{Screenshot of Wireshark log}
  \label{fig:wireshark}
\end{figure}

Wireshark displays tons of data on each packet to include packet number of the capture, source IPs, destination Ips, protocols, frame information, mac addresses and packet data.

\subsection{Snort}

Snort is another tool that can be utilized by an examiner to analyze network traffic. Snort is different from Wireshark in the fact that it can also be used as an Intrusion Protection System. Snort is a type of tool that an enterprise client would likely use such as a network administrator for a company. Snort can alert users of malicious packets set by rules and also log packets for historic analysis. 
%Below is a screenshot of the Snort interface:

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\linewidth]{imgs/snort.jpg}
  \caption{Screenshot of Snort interface}
  \label{fig:snort}
\end{figure}

\subsection{Engage}

Several tools that examiner utilize in 802.11 forensics can also be used be attackers for malicious purposes. One such tool is the Engage Packet Builder tool. The Engage Packet Builder tool allows the user to create/inject packets and most notable allows the user to be able to spoof their IP address to become anonymous. This tool can be used for DOS type attacks. 
%Below is a screenshot from the Engage Packet Builder tool’s interface:

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\linewidth]{imgs/engage.jpg}
  \caption{Screenshot of the Engage Packet Builder tool's interface}
  \label{fig:engage}
\end{figure}

\subsection{Aircrack}

Attackers can also utilize Wireshark and a tool called Aircrack-ng in order to crack a WAP’s password. An attacker can sniff out packets with a capable wireless card and capture them utilizing Wireshark. If the attacker is able to capture the correct type of packets, then they can use Aircrack-ng to brute force crack the password. Once the attacker has the encryption password then they can see the data behind all of the packets which could be sensitive data the authorized user of the network does not want to share. 
%Below is a screenshot of an example of Aircrack-ng cracking a password utilizing a brute force method:

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\linewidth]{imgs/aircrack.jpg}
  \caption{Screenshot of the Engage Packet Builder tool's interface}
  \label{fig:aircrack}
\end{figure}

\subsection{Universal Forensic Extraction Device}

“The term “mobile devices” encompasses a wide array of gadgets ranging from mobile phones, smartphones, tablets, and GPS units to wearables and PDAs” \cite{Kostadinov_2019}. Similarly to WiFi, mobile devices have become part of the everyday norm for people. A person can usually not go more than 10 minutes without checking their mobile phone unless they are asleep. Mobile devices such as cell phone can contain a ton of personal information, from banking information to passwords and because of that, mobile devices are a key in some investigations and a target for attackers. 
UFED (Universal Forensic Extraction Device) by Cellebrite is a powerful licensed tool for mobile forensics that allows an investigator to bypass pattern, password or PIN locked devices on both Android and iOS devices, collect data from phones, SIM cards, SD cards, GPS devices, and access data from a variety of different apps. 
%Below is a screenshot of one of the UFED menu lists:

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\linewidth]{imgs/ufed.jpg}
  \caption{Screenshot of the UFED menu lists}
  \label{fig:ufed}
\end{figure}

\subsection{iOS Forensic Toolkit}

For iOS specific forensics, the Elcomsoft iOS Forensic Toolkit is extremely useful. This tool can extract all of the data on an iOS device to include, backups, crash logs, media, shared files and can automatically disable screen lock once a device is connected. This tool will work with and iOS device including Apple watches and TV. %Below is a screenshot of the iOS Forensic Toolkit’s main menu:

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\linewidth]{imgs/ftk.jpg}
  \caption{Screenshot of iOS Forensic Toolkit’s main menu}
  \label{fig:ftk}
\end{figure}

\subsection{Oxygen}

Oxygen Forensic Detective is another powerful tool for mobile forensics. This tool is not only useful for cell phones, but also useful for conducting forensics on IoT devices, device backups, UICC and media cards, drone, and cloud services. It has similar capabilities to the last two mobile forensic tools discussed in that it can bypass screen locks and then extract data. 
%Below is a screenshot of Oxygen Forensic Detective:

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\linewidth]{imgs/oxygen.jpg}
  \caption{Screenshot of Oxygen Forensic Detective}
  \label{fig:oxygen}
\end{figure}

Overall, this was just a small sample size of the available tools for that can be used for 802.11 and mobile forensics. There are several different vendors that supply a variety of forensic tools that can be used for law enforcement, government agencies and if in the wrong hands can be used by attackers. As technology becomes more unadvanced and encryption methods become harder to bypass, tools will continue to have to be upgraded or developed in order to keep up with the ability to conduct successful sound evidence.

\section{Conclusion}
%% TODO add conclusion

\bibliographystyle{style/ACM-Reference-Format}
\bibliography{references}
